---
title: "SRPB traveling subjects reproducibility analysis"
author: "Brian B. Avants"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
---

```{r global options, include=FALSE}

library(knitr)
library(ANTsR)
library( ggplot2 )
require(grid)
require(gridExtra)
library( tidyr )
library("ggthemes")
library( ggthemr )
library(SimplyAgree)
library(DescTools)
library( subtyper )
library(GGally)

theme_set(
    theme_classic(base_size = 25)
)
ggthemr("flat",text_size=20,spacing=1.3)
update_geom_defaults("point", list(size = 4))

```

> "if not now, when?"


## reproducibility analysis for antspyt1w results

this documents different approaches to looking at reproducibility in numeric
summary data derived from T1w images by ANTsX methods.

the `csv` results are generated from antspyt1w version 0.4.9.

* report overall reproducibility for a given region

* report a summary of cross-site agreement for a given region

* potentially look at effect sizes (though the variables are limited for this)

* we should distill the best of these measurements to a single function and
then run on all of our derived variables --- thereby assigning specific
confidence to different measurements that we have.

* the `tarAnat` and `anatcsv` values should match.  the analysis is
then performed on whatever `roi$Description` matches the `tarAnat` `grep`.

* the Achieva appears to have the highest quality wrt `cnr` and `snr_total`.

* overall, YC1 is among the best sites and subject sub-094 may have the
best quality rating.

```{r importantParameters}
excludeScanners = c( "Signa", "Spectra" ) # eliminate the really bad scanner
excludeSite = c( "UTO" ) # eliminate the really bad scanner
excludeSite = c( "HKH" ) # eliminate the really bad scanner
excludeScanners = c(  "Signa" ) #, "MR", "Achieva" )
excludeScanners = c(  ) # eliminate the really bad scanner
excludeSite = c(  ) # eliminate the really bad scanner
onlyDayOne = TRUE # eliminate the multi-day data collected only at one site
# tarAnat = c( "vol_nbm_left_antbf", "vol_nbm_left_posbf", "vol_nbm_right_midbf",
# "vol_nbm_left_midbf", "vol_nbm_right_antbf",  "vol_nbm_right_posbf" )
# tarAnat = c("area_left_perirhinalmtl" ,  "area_right_perirhinalmtl")
# tarAnat=c("vol_mtg_sn_snc_leftcit168","vol_mtg_sn_snc_rightcit168")
# tarAnat=c("vol_mtg_sn_snc_leftdeep_cit168","vol_mtg_sn_snc_rightdeep_cit168")
# tarAnat=c("vol_mtg_sn_snc_leftsnseg","vol_mtg_sn_snc_rightsnseg")
# tarAnat=c( "vol_deepgraytissues", "vol_gmtissues", "vol_wmtissues" )
# tarAnat = c( "thk_left_hippocampushipplr", "thk_right_hippocampushipplr")
# tarAnat = c( "thk_left_hippocampushipplr", "thk_right_hippocampushipplr")
# tarAnat = c("vol_left_ca1mtl","vol_right_ca1mtl")
# tarAnat = "vol_gmtissues"
if ( ! exists( "tarAnat") ) {
  tarAnat = 'vol_mtg_sn_snc_leftsnseg'
}
if ( ! exists( "anatcsv") ) {
  anatcsv='_v2SR_mergewide.csv'
  anatcsv='_v0SR_mergewide.csv'
  # anatcsv='_v3SR_mergewide.csv'
  # anatcsv='_v1SR_mergewide.csv'
  }

post=c('SRHIER','')[2]
print(tarAnat)
```


```{r,echo=FALSE}
multigrep <- function( x, desc ) {
  roisel = c()
  for ( xx in x )
    roisel = c( roisel, grep(xx, desc) )
  return(  roisel )
}

library(ANTsR)
rdir=path.expand("~/code/antsxReproducibility")
dd=read.csv(paste0( rdir, "/SRPBTravel/participants.csv"))
qct1 = read.table( paste0( rdir, "/QC/group_T1w.tsv"),header=TRUE)
qct1$participants_id=gsub("_T1w","",qct1$bids_name)
qct1$participants_id=gsub("_run-01","",qct1$participants_id)
dd = merge(dd,qct1,by='participants_id')
usubs = unique( dd$Subject )
usite = unique( dd$Site )
uscan = unique( dd$Scanner )
undone=TRUE
for ( j in 1:nrow(dd)) {
  myid = dd[j,"participants_id"]
  fn = paste0( rdir, "/SRPBTravel/", myid,
    "/T1wHierarchical/SRPBTravel-",myid,"-T1wHierarchical-",post,anatcsv)
  fn = paste0( rdir, "/traveling_subjects_repro_study/", myid,
    "/T1wH/",myid,"_T1wH",post,anatcsv)
  if ( j==1 ) print(fn)
  if( file.exists( fn ) ) {
    roi = read.csv( fn )[,-1]
    if ( undone ) {
      bounddf = data.frame( matrix(NA,nrow=nrow(dd),ncol=ncol(roi)))
      colnames(bounddf)=colnames(roi)
      dd = cbind( dd, bounddf )
      undone = FALSE
    }
    dd[j,colnames(roi)] = roi
    }
  }
if ( length( tarAnat ) > 1 ) {
  dd$outcome = rowSums( dd[,tarAnat] )
} else   dd$outcome = ( dd[,tarAnat] )
if ( onlyDayOne ) dd = dd[ dd$Day == 1, ]
if ( length( excludeScanners ) > 0 )
  dd=dd[ ! (dd$Scanner %in% excludeScanners ),] # eliminate the really bad scanner
if ( length( excludeSite ) > 0 )
  dd=dd[ ! (dd$Site %in% excludeSite ),] # eliminate the really bad scanner
dd = dd[ !is.na( dd$RandBasisProj01 ),  ]

```


```{r ddoutlieranalysis,echo=FALSE,eval=TRUE,message=FALSE}
olcols = grep( "RandBasisProj[0-9]", colnames(dd))
oldf = dd[, olcols ]
oldf = subtyper::outlierness( oldf, calck=24 )
ddOL = cbind(dd,oldf[,grep("OL_",names(oldf))])

olnames = names(oldf)[ grep("OL_",names(oldf)) ]
for ( nm in olnames ) {
#  print(nm)
#  print(table( ddOL$Scanner, ddOL[,nm] > quantile( ddOL[,nm], 0.66 ) ))
  }

dd$OL_LOOP=ddOL$OL_LOOP
```


## table of scanners involved here

```{r,echo=FALSE}
pander::pander( table( dd$Scanner ) )
```



## Summmary reliability data: all measurements

```{r allrel,fig.width=14,fig.height=4,echo=FALSE,eval=TRUE,message=FALSE,warning=FALSE}
# now site reliability
varstostudy = names( roi )[ -c(1:4) ]
varstostudy = varstostudy[ -grep("RandBasisProj",varstostudy)]
relidf = data.frame( label=varstostudy )
relidf['mean']=NA
relidf['icc']=NA
relidf['cv']=NA
ct=1
for ( myvar in varstostudy ) {
  mypw = pivot_wider( dd, id_cols='Subject',
    names_from=c("Site"), values_from=tidyselect::all_of(myvar) )
  myiccs = ICC( mypw[,-1] )
  # pander::set.caption(paste( "Site ICC2k - average k raters"))

  temp = reli_stats( myvar, "Scanner", "Subject", data=dd )
  relidf[ct,c("mean","icc","cv")] = c(
    mean(dd[,myvar],na.rm=TRUE),
    myiccs$results[5,2],
    temp$cv )
  ct=ct+1
  }

# library(RColorBrewer)
# n = length( varstostudy )
# qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
# col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
# pie(rep(1,n), col=sample(col_vector, n,replace=TRUE))

# library(ggsci)
# ugly <- define_palette(
#      swatch = ,
#      # c('black', 'red', 'green', 'blue', 'brown', 'purple', 'yellow'),
#      gradient = c(lower = 'red', upper = 'pink')
#    )
# ggthemr(ugly)
# ggplot(relidf, aes(fill=label, y=icc, x=label)) +
#    geom_bar(position="dodge", stat="identity")+
#      theme(text = element_text(size=22),
#        axis.text.x = element_text(angle=0, hjust=1))
library(plotly)
relidf = relidf[ order( relidf$mean),  ]
relidfvol = relidf[ grep("vol_",relidf$label),]
relidfvol = relidfvol[ order( relidfvol$cv),  ]
relidfvol$label = factor( relidfvol$label, levels=relidfvol$label)
relidfarea = relidf[ grep("area_",relidf$label),]
relidfarea = relidfarea[ order( relidfarea$cv),  ]
relidfarea$label = factor( relidfarea$label, levels=relidfarea$label)
relidfthk = relidf[ grep("thk_",relidf$label),]
relidfthk = relidfthk[ order( relidfthk$cv),  ]
relidfthk$label = factor( relidfthk$label, levels=relidfthk$label)

plot_ly(relidfvol, type='bar', x = ~label, y = ~cv, text = ~label, name="",
               texttemplate = '%{y:.2s}', textposition = 'outside') %>% layout(uniformtext=list(minsize=8, mode='hide'))
#
plot_ly(relidfarea, type='bar', x = ~label, y = ~cv, text = ~label, name="",
               texttemplate = '%{y:.2s}', textposition = 'outside') %>% layout(uniformtext=list(minsize=8, mode='hide'))

#
plot_ly(relidfthk, type='bar', x = ~label, y = ~cv, text = ~label, name="",
               texttemplate = '%{y:.2s}', textposition = 'outside') %>% layout(uniformtext=list(minsize=8, mode='hide'))


plot_ly(relidfvol, x = ~mean, y = ~icc, type = 'scatter', mode = 'markers',
        text = ~paste('Vol: ', label))

fig = plot_ly(relidfvol[ relidfvol$mean<quantile(relidfvol$mean,0.2), ], x = ~mean, y = ~icc, type = 'scatter', mode = 'markers', text = ~paste('Vol: ', label))
subsel = multigrep( c("bf","snr","snc"), relidfvol$label)
fig <- fig %>% add_annotations(x = relidfvol$mean[subsel],
                          y = relidfvol$icc[subsel],
                          text = relidfvol$label[subsel],
                          xref = "x",
                          yref = "y",
                          showarrow = TRUE,
                          arrowhead = 4,
                          arrowsize = .5,
                          ax = 20,
                          ay = -40)

#
volsel=relidfvol$mean>quantile(relidfvol$mean,0.5) &
 relidfvol$mean<quantile(relidfvol$mean,0.66)
relisub = relidfvol[ volsel, ]
fig = plot_ly(relisub, x = ~mean, y = ~icc, type = 'scatter', mode = 'markers', text = ~paste('Vol: ', label))
subsel = multigrep( c("vol_bn_str_pu","vol_bn_str_ca"), relisub$label)
fig %>% add_annotations(x = relisub$mean[subsel],
                          y = relisub$icc[subsel],
                          text = relisub$label[subsel],
                          xref = "x",
                          yref = "y",
                          showarrow = TRUE,
                          arrowhead = 4,
                          arrowsize = .5,
                          ax = 20,
                          ay = -40)

plot_ly(relidfthk, x = ~mean, y = ~icc, type = 'scatter', mode = 'markers',
        text = ~paste('Thk: ', label))

#
plot_ly(relidfarea, x = ~mean, y = ~icc, type = 'scatter', mode = 'markers',
        text = ~paste('Area: ', label))
plot_ly(relidfarea[ relidfarea$mean<quantile(relidfarea$mean,0.2), ], x = ~mean, y = ~icc, type = 'scatter', mode = 'markers',  text = ~paste('Area: ', label))

```


## show volumes for one subject at different sites

```{r onesub3bars,fig.width=17,fig.height=6,echo=FALSE}
ggthemr("flat")
ggthemr('pale', layout = 'scientific', spacing = 2, type = 'inner')
nvols = length( tarAnat )
dfvols = dd[which( dd$Subject == "MP004" ),
  c("Subject","participants_id","Site", "Scanner",tarAnat)]
dfvols = dfvols[ !is.na(dfvols[,tarAnat[1]]),]
dflong = pivot_longer( dfvols, tail(names(dfvols),nvols) )
dflong$Subject=factor(dflong$Subject)
names(dflong)=c("ID", "image", "Site", "Scanner","Anat","Volume")
dflong$Volume = as.numeric( dflong$Volume )
ylims = c( median(dflong$Volume ), max(dflong$Volume)+0.05*max(dflong$Volume))
ggplot(dflong, aes(fill=Scanner, y=Volume, x=Anat)) +
    geom_bar(position="dodge", stat="identity")+
      theme(text = element_text(size=22),
        axis.text.x = element_text(angle=0, hjust=1))
#
#ggplot(dflong, aes(fill=Site, y=Volume, x=Anat)) +
#    geom_bar(position="dodge", stat="identity")+
#      theme(text = element_text(size=22),
#        axis.text.x = element_text(angle=0, hjust=1))

```


## Summmary  reliability data: statistical measures of reproducibility

```{r,fig.width=14,fig.height=4,echo=FALSE}
# now site reliability
mypw = pivot_wider( dd, id_cols='Subject',
  names_from=c("Site"), values_from='outcome' )
myiccs = ICC( mypw[,-1] )
pander::set.caption(paste( "Site ICC2k - average k raters"))
pander::pander( myiccs$results[5,1:6] )

print( reli_stats( "outcome", "Scanner", "Subject", data=dd ) )
plot( reli_stats( "outcome", "Scanner", "Subject", data=dd ) )

```



## Demonstrate effect of subject vs effect of Scanner or Site: Raw data plots

```{r plotbysubject,fig.width=12,fig.height=12,eval=FALSE,echo=FALSE,message=FALSE}
mycols = c('age', 'outcome', 'Site', 'Day','Phase','Manufacture','Coil','Scanner')
mycols = c('age', 'outcome', 'Site', 'Scanner')
nna = ! is.na( dd$outcome )
print(
  ggpairs(data = dd[nna, c('Subject',mycols)] ) )
#    mapping = ggplot2::aes_string(color = 'Subject'),
#        columns = mycols,
#        upper = list(continuous = "cor",
#                     combo = "box_no_facet",
#                     discrete = "facetbar",
#                     na = "na"),
#        lower = list(continuous = "points",
#                     combo = "facethist",
#                     discrete = "facetbar",
#                     na = "na"),
#        diag = list(continuous = "densityDiag",
#                    discrete = "barDiag",
#                    na = "naDiag"),
#        xlab = 'X',
#        ylab = 'Y',
#    title = 'Exploratory correlations:')


```

```{r gg2,fig.width=16,fig.height=5,echo=FALSE}
ggthemr("fresh",text_size=20,spacing=1.3)
  ggplot(dd, aes(x = Subject, y = outcome, color = Site)) +
  geom_jitter(width=0.2,height=0.2) + ggtitle("Site")
```


```{r gg2site,fig.width=16,fig.height=5,echo=FALSE}
ggplot(dd, aes(x = Subject, y = outcome, color = Scanner)) +
  geom_jitter(width=0.2,height=0.2) + ggtitle("Scanner") #+geom_point(size=3)
```


## Demonstrate effects of controlling for different variables with regression



What is the fit of the model if we predict with subject alone?
```{r statsage,fig.width=10,fig.height=4}
mdl=(lm( outcome ~  1 + Subject + 1 , data=dd ))
rsq2 = format(summary( mdl )$r.squared,digits=3)
visreg::visreg( mdl, "Subject", gg=TRUE ) + ggtitle(paste("Subject only R^2:",rsq2))
```

What is the variability of the measurement if we add manufacturer?
```{r statsageb,fig.width=10,fig.height=4}
if ( length( unique(dd$Manufacture)) > 1 ) {
  mdl=(lm( outcome ~  Manufacture + Subject + OL_LOOP , data=dd ))
  rsq2 = format(summary( mdl )$r.squared,digits=3)
  visreg::visreg( mdl, "Subject", gg=TRUE ) + ggtitle(paste("Subject + Mfg R^2:",rsq2))
  }
```

What is the reproducibility of the measurement if add scanner?
```{r stats,fig.width=10,fig.height=4}

mdl=(lm( outcome ~   Subject + OL_LOOP + Scanner , data=dd ))
rsq2 = format(summary( mdl )$r.squared,digits=3)
visreg::visreg( mdl, "Subject", gg=TRUE ) + ggtitle(paste("Subject + scanner R^2:",rsq2))
```

What is the accuracy of the model if we add site?
```{r stats2c,fig.width=10,fig.height=4}
mdl=(lm( outcome ~   Subject + OL_LOOP + Site, data=dd ))
rsq2 = format(summary( mdl )$r.squared,digits=3)
visreg::visreg( mdl, "Subject", gg=TRUE ) + ggtitle(paste("Subject + site R^2:",rsq2))
```


What is the effect of total SNR?
```{r stats2b,fig.width=10,fig.height=4}

mdl=(lm( outcome ~   Subject + OL_LOOP + Scanner + snr_total , data=dd ))
rsq2 = format(summary( mdl )$r.squared,digits=3)
visreg::visreg( mdl, 'snr_total', gg=TRUE )+ ggtitle(paste("Subject + scanner + snr R^2:",rsq2))
```


What is the effect of gray matter SNR?
```{r stats3,fig.width=10,fig.height=4}

mdl=(lm( outcome ~   Subject + OL_LOOP + Scanner + snr_gm , data=dd ))
rsq2 = format(summary( mdl )$r.squared,digits=3)
visreg::visreg( mdl, 'snr_gm', gg=TRUE )+ ggtitle(paste("Subject + scanner + gm snr R^2:",rsq2))

```



What is the effect of motion confounds?

see the discussion of MRIQC [here](https://cpb-us-w2.wpmucdn.com/sites.brown.edu/dist/0/278/files/2021/10/MRIQC.pdf)


```{r statsnorun,echo=FALSE,eval=FALSE,fig.width=10,fig.height=4}
qnms=names(dd)[13:80]
for ( qnm in qnms ) {
  mdl=(lm( paste0("outcome ~   Subject + Scanner + ", qnm) , data=dd ))
  mycoffs=coefficients(summary(mdl))
  if ( mycoffs[qnm,4] < 0.01 )  {
    print(mycoffs)
    print(visreg::visreg( mdl, qnm, gg=TRUE ))
    Sys.sleep( 1 )
  }
}
```


What is the reproducibility of the measurement if we control for a motion specific variable?

```{r stats2,fig.width=10,fig.height=4}

mdl=(lm( outcome ~   Subject + OL_LOOP + Scanner + efc , data=dd ))
rsq2 = format(summary( mdl )$r.squared,digits=3)
visreg::visreg( mdl, 'efc', gg=TRUE )+ ggtitle(paste("Subject + scanner + efc R^2:",rsq2))

# grid.arrange( grobs = visreg::visreg( mdl, gg=TRUE ), ncol=1, main='reproducibility' )
```

##  Site-wise reliability

```{r,fig.width=12,fig.height=12,echo=FALSE}
library(irr)
myvar='Site'
# myvar='Scanner'
uscans=unique(dd[,myvar])
mytable = table( dd[,myvar] )
uscans = names(mytable[mytable > 2 ])
nVars=length(uscans)
iccdf = matrix(0,nrow=nVars,ncol=nVars)
uscansNam=uscans
for ( j in 1:nVars ) {
  ddscanj = dd[dd[,myvar]==uscans[j] & dd$Day==1,]
  uscansNam[j]=paste0(uscansNam[j],unique(ddscanj$Scanner))
}
rownames(iccdf)=uscansNam
colnames(iccdf)=uscansNam
for ( k in 1:nVars) {
  for (j in 1:nVars) {
  if ( TRUE) {
    subsloc =  intersect(
      dd[dd[,myvar]==uscans[k],"Subject"],
      dd[dd[,myvar]==uscans[j],"Subject"] )
    ddscanj = dd[dd[,myvar]==uscans[j] & dd$Day==1,]
    ddscank = dd[dd[,myvar]==uscans[k] & dd$Day==1,]
    sdf=data.frame(
      V0=ddscank[ ddscank[,"Subject"] %in% subsloc,"outcome"],
      V1=ddscanj[ ddscanj[,"Subject"] %in% subsloc,"outcome"] )
    iccSR = irr::icc( sdf[,c("V0","V1")],
      model = "oneway", type = "agreement", unit = "single" )$value # ,1,5)
    iccdf[k,j]=iccSR
    }
  }}#
# pheatmap::pheatmap(iccdf,display_numbers=T)
library(corrplot)
corrplot.mixed( ( iccdf ),is.corr=F)
# corrplot::corrplot(iccdf, method = 'shade', order = 'AOE', diag = FALSE)
library("gplots")
heatmap.2( ( iccdf ), scale = "none", col = bluered(100),
          trace = "none", density.info = "none")
```



```{r,echo=FALSE,eval=FALSE}
achievaPIDs=dd$participants_id[ dd$Scanner == "Achieva" ]
achievaSubs=dd$Subject[ dd$Scanner == "Achieva" ]
achievaSites=dd$Site[ dd$Scanner == "Achieva" ]
usubs = unique( achievaSubs )
bestSIDsYC1 = dd$Subject[ dd$Site == "YC1" ]
bestSIDsHUH = dd$Subject[ dd$Site == "HUH" ]
bestPIDsYC1 = dd$participants_id[ dd$Site == "YC1" ]
bestPIDsHUH = dd$participants_id[ dd$Site == "HUH" ]

# form a data frame with the necessary mappings
for ( k in 1:nrow( dd ) ) {
  sid = dd$Subject[ k ]
  if ( sid %in% bestSIDsYC1 ) {
    gtid = bestPIDsYC1[  which(bestSIDsYC1 == sid ) ]
    gtidk = which( dd$participants_id == gtid )
  } else if ( sid %in% bestSIDsHUH ) {
    gtid = bestPIDsHUH[  which(bestSIDsHUH == sid ) ]
    gtidk = which( dd$participants_id == gtid )
    }
  # get the images
#  ./SRPBTravel/sub-042/T1wHierarchical/SRPBTravel-sub-042-T1wHierarchical-SRHIERbrain_n4_dnz.nii.gz
  exts = c("-SRHIERbrain_n4_dnz.nii.gz","-cit168lab.nii.gz","-bf.nii.gz")
  gtfn = paste0( rdir, "/SRPBTravel/", dd$participants_id[gtidk], "/T1wHierarchical/SRPBTravel-", dd$participants_id[gtidk], "-T1wHierarchical", exts )
  gtImage = antsImageRead( gtfn[1] )
  gtImageCIT = antsImageRead( gtfn[2] )
  gtImageBF = antsImageRead( gtfn[3] )
  tarfn = paste0( rdir, "/SRPBTravel/", dd$participants_id[k], "/T1wHierarchical/SRPBTravel-", dd$participants_id[k], "-T1wHierarchical", exts )
  targetImage = antsImageRead( tarfn[1] )
  targetImageCIT = antsImageRead( tarfn[2] )
  targetImageBF = antsImageRead( tarfn[3] )
  temp=histogramMatchImage( targetImage, gtImage,
    numberOfHistogramBins = 48, numberOfMatchPoints=12 )
  rigidreg = antsRegistration( temp, gtImage, 'Rigid' )
  citmap = antsApplyTransforms( targetImage, gtImageCIT, rigidreg$fwdtransforms, interpolator='genericLabel')
  bfmap = antsApplyTransforms( targetImage, gtImageBF, rigidreg$fwdtransforms, interpolator='genericLabel')
  outfn = paste0( "SRPBtrainingData/", dd$participants_id[k], exts )
  antsImageWrite( targetImage, outfn[1] )
  antsImageWrite( citmap, outfn[2] )
  antsImageWrite( targetImage, outfn[3] )
  }

```
